---
title: "Hands_onExercise4_2"
author: "Rajesh babu Santigari"
format: html
editor: visual
---
#Visual Statistical Analysis
##Visual Statistical Analysis with ggstatsplot
###  ggstatsplot  is an extension of ggplot2 package for creating graphics with details from statistical tests included in the information-rich plots themselves. ###

##Getting Started
###Installing and launching R packages
```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

###Importing data

```{r}
exam <- read_csv("C:\\RBSantigari-MITB\\ISSS608-VAA\\HandsonExercise\\Hands_onExercise4(2)\\data\\Exam_data.csv")
```

###One-sample test: gghistostats() method

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

###Unpacking the Bayes Factor

```{r}

# Fit the null and alternative models
null_model <- lm(ENGLISH ~ 1, data = exam)          # H0: ENGLISH not affected by RACE
alt_model  <- lm(ENGLISH ~ RACE, data = exam)       # H1: ENGLISH affected by RACE

# Compute BICs
bic_null <- BIC(null_model)
bic_alt  <- BIC(alt_model)

# Compute Bayes Factor (BF10)
bf_10 <- exp((bic_null - bic_alt) / 2)

# Display results
cat("BIC (Null Model):", bic_null, "\n")
cat("BIC (Alternative Model):", bic_alt, "\n")
cat("Bayes Factor (BF10):", bf_10, "\n")

```
Interpretation of above results:
BF₁₀ ≈ 3009.14 → Very strong evidence in favor of the alternative hypothesis.

This suggests that RACE has a significant effect on ENGLISH scores in your data.

###Two-sample mean test: ggbetweenstats()

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

###Oneway ANOVA Test: ggbetweenstats() method

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

###Significant Test of Correlation: ggscatterstats()

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

###Significant Test of Association (Depedence) : ggbarstats() methods

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

####In this code chunk below ggbarstats() is used to build a visual for Significant Test of Association

```{r}
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```


