---
title: "Hands_onExercise2"
author: "Rajesh babu Santigari"
format: html
editor: visual
---

# Load required packages including extrafont
```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, extrafont)
```

# Load fonts into memory (important for hrbrthemes)
```{r}
extrafont::loadfonts(device = "win")
```

#2  Beyond ggplot2 Fundamentals
##2.2 Getting started
###2.2.1 Installing and loading the required libraries
```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```
###2.2.2 Importing data
```{r}
exam_data <- read_csv("C:\\RBSantigari-MITB\\ISSS608-VAA\\HandsonExercise\\Hands_onExercise2\\data\\Exam_data (1).csv")
```
##2.3 Beyond ggplot2 Annotation: ggrepel
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
###2.3.1 Working with ggrepel
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
##2.4 Beyond ggplot2 Themes
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_linedraw()+
  ggtitle("Distribution of Maths scores") 
```
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_classic() +
  ggtitle("Distribution of Maths scores") 
```
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_dark() +
  ggtitle("Distribution of Maths scores") 
```
###2.4.1 Working with ggtheme package
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
###2.4.2 Working with hrbthems package
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```
##2.5 Beyond Single Graph
## 2.5 Beyond Single Graph
###Distribution of Maths score
```{r}
p1 <- ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") + 
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of Maths scores")
print(p1)
```
###Distribution of English score
```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
print(p2)
```
###scatterplot for English score versus Maths score
```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
print(p3)
```
###2.5.1 Creating Composite Graphics: pathwork methods
###2.5.2 Combining two ggplot2 graphs
```{r}
p1 + p2
```
###2.5.3 Combining three ggplot2 graphs
```{r}
(p1 / p2) | p3
```
###2.5.4 Creating a composite figure with tag
```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```
###2.5.5 Creating figure with insert
```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```
###2.5.6 Creating a composite figure by using patchwork and ggtheme
```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```


# Extra work
##(1)ggrepel provides geoms for ggplot2 to repel overlapping text labels:
```{r}
library(ggrepel)
ggplot(exam_data, aes(MATHS, GENDER, label = rownames(exam_data))) +
  geom_text_repel() +
  geom_point(color = 'red') +
  theme_classic(base_size = 16)
```

###(2) plot for top scoring students
```{r}
library(ggrepel)

exam_data %>%
  ggplot(aes(x = MATHS, y = ENGLISH)) +
  geom_point(color = "steelblue") +
  geom_text_repel(aes(label = ifelse(MATHS > 90 | ENGLISH > 90, ID, "")), 
                  size = 3, box.padding = 0.3) +
  ggtitle("Top Scoring Students in Maths or English")
```
####(3) Combine plots
```{r}
library(patchwork)

p1 <- ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, fill = "lightblue") +
  ggtitle("Maths")

p2 <- ggplot(exam_data, aes(x = ENGLISH)) +
  geom_histogram(bins = 20, fill = "lightgreen") +
  ggtitle("English")

p1 + p2 
```
###(4)English Score by Class and Gender
```{r}
ggplot(exam_data, aes(x = CLASS, y = ENGLISH, fill = GENDER)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +
  ggtitle("English Score by Class and Gender")
```
###(5)Compare subject-wise score distributions in a single view
```{r}
library(patchwork)

p1 <- ggplot(exam_data, aes(x = ENGLISH)) + 
  geom_histogram(bins = 20, fill = "steelblue") +
  ggtitle("English Scores")

p2 <- ggplot(exam_data, aes(x = MATHS)) + 
  geom_histogram(bins = 20, fill = "tomato") +
  ggtitle("Maths Scores")

p3 <- ggplot(exam_data, aes(x = SCIENCE)) + 
  geom_histogram(bins = 20, fill = "darkgreen") +
  ggtitle("Science Scores")

(p1 | p2) / p3
```
